The Name of the Title Is Hope

### 2025-01-28 paper.tex
Add Graphs (Placeholder Data)

\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{listings}
\newcommand{\thomas}[1]{{\footnotesize\color{orange}[Thomas: #1]}}
\newcommand{\john}[1]{{\footnotesize\color{cyan}[John: #1]}}
\newcommand{\raph}[1]{{\footnotesize\color{magenta}[Raph: #1]}}

@@

The viability of \emph{Decoupled-Fallback} relies on the same cache locality properties as \emph{Decoupled-Lookback}. No subgroup scheduler is perfectly fair, and thus any given hardware will have a natural fallback-incidence rate $f$, regardless of FPG, due to \emph{false-positive} fallback operations. Consequently, we expect memory bandwidth traffic to increase (and throughput to decrease) by $O(fn)$ to $O((2 + f)n)$. However, just as serial ordering guarantees that memory accesses to the spine fall within an $o$-width sliding window, memory accesses to the scan input are similarly guaranteed to fall within an $ot$-width sliding window, making them also highly likely to reside within some level of the cache. As we later empirically demonstrate, this cache residency significantly reduces the cost of memory access, such that the fallback operation is almost always bound by the compute required to reduce the blocking tile. As a pure reduction incurs very lower overhead, we observe throughput approaching $O(2n)$ on most devices.

@@

  \item \textbf{Fallback:}
        \begin{itemize}
          \item The workgroup computes the reduction of the blocking tile by processing the corresponding elements directly.
          \item Once the fallback reduction is complete, the workgroup \emph{attempts} to post the reduction to global memory and update the tile's state to \emph{Ready}.
          \item As this update attempt is performed atomically, the workgroup reads back the blocking tile's immediate state:
                \begin{itemize}
                  \item If the state is \emph{Inclusive}, the fallback value is discarded, and the workgroup accumulates the value before proceeding to step 4.
                  \item If the blocking tile is the first in the spine, the workgroup has reached the start of the spine and proceeds to step 4.
                  \item Otherwise, the workgroup resumes the lookback, starting with the preceding tile.
                \end{itemize}
        \end{itemize}

  \item \textbf{Finalization:}
  \begin{itemize}
      \item Once the lookback or fallback is complete, the workgroup posts its inclusive reduction to global memory, transitioning its tile state to \emph{Inclusive}.
      \item The \emph{previous reduction} is also posted to shared memory for the workgroup to consume and propagate as necessary.
      \item The control flag is updated to \emph{Unlocked} to signal the exit of the \emph{Decoupled-Fallback} phase, allowing subsequent operations to proceed.
  \end{itemize}
\end{enumerate}

@@

In the context of fallbacks, we require a tile state representation and an accompanying atomic operation that ensures a fallback update attempt can safely transition a tile's state to \emph{Ready} without overwriting a more advanced state. When an update attempt is made, the blocking tile can exist in any of the three states due to two scenarios: a false-positive fallback may have been initiated on a tile that transitions to \emph{Ready} or \emph{Inclusive} before the fallback completes, or another workgroup may have already successfully completed a fallback update, leaving the tile in \emph{Ready}.Consequently, the update mechanism should only overwrite tiles in the \emph{Not Ready} state. This requirement is further constrained by the fact that in WebGPU, the correctness of atomic \emph{Compare-and-Swap} is not guaranteed across all devices, prohibiting its use. To address this, we encode the tile state within the most significant bits of the split value and enforce a strict state ordering: \emph{Inclusive} $>$ \emph{Ready} $>$ \emph{Not Ready}. This design enables the use of atomic \emph{Maximum}, which inherently respects the state hierarchy, to safely perform updates.

\section{Evaluation}
256 threads, 16 elements per thread. Shared memory requirements. RTS implementation shares an identical local implementation. Uses Merrill-Grimshaw style raking, with our own occupancy discovery technique.

\subsection{Comparison to Reduce-Then-Scan}

@@

\subsection{Emulated Deadlocking}
Show the performance of devices as we artificially increase the fallback-incidence rate.

>>>>

\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{listings}
\usepackage{subcaption}
\newcommand{\thomas}[1]{{\footnotesize\color{orange}[Thomas: #1]}}
\newcommand{\john}[1]{{\footnotesize\color{cyan}[John: #1]}}
\newcommand{\raph}[1]{{\footnotesize\color{magenta}[Raph: #1]}}

@@

The viability of \emph{Decoupled-Fallback} relies on the same cache locality properties as \emph{Decoupled-Lookback}. No subgroup scheduler is perfectly fair, and thus any given hardware will have a natural fallback rate $f$, regardless of FPG, due to \emph{false-positive} fallback operations. Consequently, we expect memory bandwidth traffic to increase (and throughput to decrease) by $O(fn)$ to $O((2 + f)n)$. However, just as serial ordering guarantees that memory accesses to the spine fall within an $o$-width sliding window, memory accesses to the scan input are similarly guaranteed to fall within an $ot$-width sliding window, making them also highly likely to reside within some level of the cache. As we later empirically demonstrate, this cache residency significantly reduces the cost of memory access, such that the fallback operation is almost always bound by the compute required to reduce the blocking tile. As a pure reduction incurs very lower overhead, we observe throughput approaching $O(2n)$ on most devices.

@@

  \item \textbf{Fallback:}
        \begin{itemize}
          \item The workgroup computes the reduction of the blocking tile by processing the corresponding elements directly. This process involves a workgroup-wide reduction using additional shared memory to aggregate partial subgroup results.
          \item Once the fallback reduction is complete, the workgroup \emph{attempts} to post the reduction to global memory and update the tile's state to \emph{Ready}.
          \item As this update attempt is performed atomically, the workgroup reads back the blocking tile's immediate state:
                \begin{itemize}
                  \item If the state is \emph{Inclusive}, the fallback value is discarded, and the workgroup accumulates the value before proceeding to step 4.
                  \item If the blocking tile is the first in the spine, the workgroup has reached the start of the spine and proceeds to step 4.
                  \item Otherwise, the workgroup resumes the lookback, starting with the preceding tile.
                \end{itemize}
        \end{itemize}

  \item \textbf{Finalization:}
  \begin{itemize}
      \item Upon completing the lookback or fallback, the local reduction is added to the \emph{previous reduction} to compute the inclusive reduction, which is then posted to global memory, transitioning the tile state to \emph{Inclusive}.
      \item The \emph{previous reduction} is posted to shared memory for the workgroup to consume and propagate as necessary.
      \item The control flag is updated to \emph{Unlocked} to signal the exit of the \emph{Decoupled-Fallback} phase, allowing subsequent operations to proceed.
  \end{itemize}
\end{enumerate}

@@

In the context of fallbacks, we require a tile state representation and an accompanying atomic operation that ensures a fallback update attempt can safely transition a tile's state to \emph{Ready} without overwriting a more advanced state. When an update attempt is made, the blocking tile can exist in any of the three states due to two scenarios: a false-positive fallback may have been initiated on a tile that transitions to \emph{Ready} or \emph{Inclusive} before the fallback completes, or another workgroup may have already successfully completed a fallback update, leaving the tile in \emph{Ready}. Consequently, the update mechanism should only overwrite tiles in the \emph{Not Ready} state. This requirement is further constrained by the fact that in WebGPU, the correctness of atomic \emph{Compare-and-Swap} is not guaranteed across all devices, prohibiting its use. To address this, we encode the tile state within the most significant bits of the split value and enforce a strict state ordering: \emph{Inclusive} $>$ \emph{Ready} $>$ \emph{Not Ready}. This design enables the use of atomic \emph{Maximum}, which inherently respects the state hierarchy, to safely perform updates.

\section{Evaluation}
\subsection{Experimental Setup}
We configure workgroups with a size of 256 threads, where each thread processes 4 vectors of size 4, resulting in 16 elements per thread and a total work tile size of 4096. We find a maximum spin count of 4 as offering good responsiveness to blocking, while minimizing false-positives fallbacks. As the size of the local spine is inversely proportional to the subgroup size, we allocate sufficient shared memory to accommodate the minimum subgroup size specified by the WebGPU standard, which is 4. Despite the additional shared memory required for conflict avoidance and fallback operations, total shared memory usage remains modest at only 1 KB. For all tests, we use an input size of $2^{25}$.

Something about devices here.
\subsection{Performance Comparison}
To evaluate the performance of \emph{Decoupled-Fallback}, we compare it against two key benchmarks: \emph{Reduce-Then-Scan} (RTS) and a \emph{Memcopy} operation. RTS serves as a baseline to demonstrate the performance gains of our approach over the previous FPG-constrained technique. Memcopy is included as a point of reference, as it shares the same $2n$ global memory movement required by scan operations.
\begin{figure}[h!]
  \centering
  \includegraphics[width=\linewidth]{graphics/speedup.pdf}
\end{figure}
Since WebGPU limits device-side timestamps to compute or render passes, we implement a custom Memcopy kernel for timing purposes. Similarly, as there are no existing WebGPU implementations of RTS, we rely on our own optimized version for comparison. To ensure a competitive evaluation, our RTS implementation employs a variable-launch main kernel with a single workgroup performing a serial spine scan, as we found raking to be less performant in practice. Apart from this adjustment, the intraworkgroup implementation in RTS is identical to that used in \emph{Decoupled-Fallback}.

\thomas{Performance is good}

@@

\subsection{Simulated Blocking}
To demonstrate the resilience of our technique under fallback rates significantly higher than those observed organically on devices, we intentionally force fallbacks by omitting the posting of work tile reductions. Starting with infrequent omissions---1 in 512 workgroups---we progressively increase the rate to near-continuous levels, reaching 1 in 2 workgroups.
\begin{figure}[htbp]
  \centering
  \begin{subfigure}{.9\linewidth}
    \centering
    \includegraphics[width=\linewidth]{graphics/fallbacksInitiated_plot.pdf}
    \caption{Fallbacks Initiated}
    \label{fig:fallbacks_initiated}
  \end{subfigure}
  \begin{subfigure}{.9\linewidth}
    \centering
    \includegraphics[width=\linewidth]{graphics/totalSpins_plot.pdf}
    \caption{Total Spins}
    \label{fig:total_spins}
  \end{subfigure}
  \begin{subfigure}{.9\linewidth}
    \centering
    \includegraphics[width=\linewidth]{graphics/lookbackLength_plot.pdf}
    \caption{Lookback Length}
    \label{fig:lookback_length}
  \end{subfigure}
  \begin{subfigure}{.9\linewidth}
    \centering
    \includegraphics[width=\linewidth]{graphics/time_plot.pdf}
    \caption{Speed vs Memcopy}
    \label{fig:execution_time}
  \end{subfigure}
  \caption{Simulated Blocking}
  \label{fig:vertical_images}
\end{figure}

In Figure~\ref{fig:fallbacks_initiated}, we measure the number of fallbacks proportional to the number of workgroups launched, and verify that we are indeed inducing fallbacks. In Figure~\ref{fig:total_spins} we measure \thomas{this should be spins per workgroup it's a better metric}. In Figure~\ref{fig:lookbackLength} we see that. Finally, in Figure~\ref{fig:execution_time}, we can see that, even when \emph{half} the workgroups are deadlocking, \emph{Decoupled-Fallback} still achieves speeds of . . .
